`flippant` -- An `R` package for the automated  analysis of dithionite

### 2015-09-18 inst/manuscript/manuscript.rmd
Keep working on the data analysis flow section;

#Input Data and Data Processing
To analyse raw data stemming from dithionite scramblase assays, `flippant` 
requires for each data point information as the path to the 
fluorometer-generated spectrum, the amount of protein used to reconstitute 
proteoliposomes, the volume used in the assay before and after addition of 
dithionite (to calculate a volume correction factor), the amount of lipids 
present (to calculate a protein to phospholipid ratio or PPR) etc. The user 
supplies this information in one row of a spreadsheet (see below for supporting functionality).

Spectral input data is read by `flippant` using the information provided in the
input template. Raw spectra produced by QuantaMaster fluorometer instruments 
(Photon Technology International, Inc., Edison, New Jersey) running software 
versions FelixGX v4.1 and Felix32 v1.20, respectively, as well as a simple 
generic tab-delimited spectrum representing file format with time (s) and 
fluorescence intensity representing columns are currently supported. 
File type determination is handled algorythmically.

Based on @menon_opsin_2011 and @goren_constitutive_2014, data is processed as 
follows. Input is format checked and defaults are injected for facultative 
parameters/columns as appropriate. Fluorescence spectra are parsed, including 
determination of when dithionite was added to the sample using peak detection by
continuous wavelet transformation as implemented by the `wmtsa` `R` package 
[@percival_wavelet_2000;@constantine_wmtsa:_2013] and resetting the acquisition 
time accordingly. The zero time point henceforth corresponds to the time of 
addition. 
Pre-dithionite-addition baseline fluorescence $F_{\text{baseline}}$ is 
determined for each spectrum by averaging (median) over the ten values preceding
dithionite addition. Post-dithinonite-addition minimum fluorescence 
$F_{\text{min}}$ is analogously calculated from the last ten datapoints after 
the default 400 s or the timepoint of measurement supplied by the user in the 
parameter file. $F_{\text{min}}$ is volume-corrected based on supplied reaction 
volumes
with and without dithionite and for each spectrum/datapoint a fluorescence 
reduction is calculated as follows:
$$
y_{\text{raw}} = 1 - \frac{F_{\text{min}}}{F_{\text{baseline}}}
$$
<!--
Data are split for parallel treatment using a combined Experimental Series/Experiment identifier (see above).-->
A relative fluorescence reduction $y$ is calculated scaled to a (required) liposomes-only/no-protein control and the probability $P_{\geq 1}$ for a 
liposome to contain at least one scramblase molecules is calculated using
$$
P_{\geq 1} = \frac{(y - y_0)}{(y_{\text{max}} - y_0)}
$$
where $y$ is the relative fluorescence reduction for the proteoliposome 
measurement and $y_0$ for a measurement of protein-free liposomes. 
$y_{\text{max}}$ refers to either the maximal $y$ in the experiment or is 
derived from a monoexponential fit to the data (default) as a precaution for the 
case where the protein/phospholipid titration did not reach the plateau of the 
saturation curve. The behavior is adjustable using the `scaleTo` parameter of 
the corresponding functions (see below). Next a monoexponential curve is fitted
to the realtionship of $P_{\geq 1}$ to the protein to phospholipid ratio (PPR; 
(mg/mmol)), providing the scramblase-characterizing fit factor $\alpha$.

>>>>

#Input Data and Data Processing
To analyse raw data stemming from dithionite scramblase assays, `flippant` 
requires for each data point information on the path to the 
fluorometer-generated spectrum, the amount of protein used to reconstitute 
proteoliposomes, the assay volume before and after addition of 
dithionite (to calculate a volume correction factor), the amount of lipids 
present (to calculate a protein to phospholipid ratio or PPR) etc. The user 
supplies this information in a spreadsheet (see below for supporting 
functionality).

Spectral input data is read by `flippant` using the file paths provided in the
spreadsheet. Raw spectra produced by QuantaMaster fluorometer instruments 
(Photon Technology International, Inc., Edison, New Jersey) running software 
versions FelixGX v4.1 and Felix32 v1.20, respectively, as well as a simple 
generic tab-delimited spectrum representing file format with time (in seconds) 
and fluorescence intensity representing columns are currently supported. 
File type determination is handled algorythmically.

Based on @menon_opsin_2011 and @goren_constitutive_2014, data is processed as 
follows. Input is format checked and defaults are injected for facultative 
parameters/columns as appropriate. Fluorescence spectra are parsed. This includes 
algorithmic determination of when dithionite was added to the sample using peak
detection by continuous wavelet transformation as implemented in the 
`wmtsa` `R` package [@percival_wavelet_2000;@constantine_wmtsa:_2013]. The 
acquisition time is reset such that the zero time point henceforth corresponds 
to the time of addition. 
Pre-dithionite-addition baseline fluorescence $F_{\text{baseline}}$ is 
determined for each spectrum by averaging (median) over the ten values preceding
dithionite addition. Post-dithinonite-addition minimum fluorescence 
$F_{\text{min}}$ is analogously calculated from the last ten datapoints after 
the default 400 s or the timepoint of measurement supplied by the user in the 
parameter spreadsheet. $F_{\text{min}}$ is volume-corrected based on supplied 
reaction volumes with and without dithionite and for each spectrum/datapoint a 
fluorescence reduction is calculated as follows:
$$
y_{\text{raw}} = 1 - \frac{F_{\text{min}}}{F_{\text{baseline}}}
$$
A relative fluorescence reduction $y$ is calculated scaled to a (required) liposomes-only/no-protein control and the probability $P_{\geq 1}$ for a 
liposome to contain at least one scramblase molecules is calculated using
$$
P_{\geq 1} = \frac{(y - y_0)}{(y_{\text{max}} - y_0)}
$$
where $y$ is the relative fluorescence reduction for the proteoliposome 
measurement and $y_0$ for a measurement of protein-free liposomes. 
$y_{\text{max}}$ refers to either the maximal $y$ in the experiment or is 
derived from a monoexponential fit to the data (default) as a precaution for the 
case where the protein/phospholipid titration did not reach the plateau of the 
saturation curve. The behavior is adjustable using the `scaleTo` parameter of 
the corresponding functions (see below). Next a monoexponential curve is fitted
to the realtionship of $P_{\geq 1}$ to the protein to phospholipid ratio (PPR; 
(mg/mmol)), providing the scramblase-characterizing fit factor $\alpha$.
